---
title: "Activity Quality"
author: "Victor Corona"
date: "March 2017"
output: html_document
---

# Background

This analysis aims at identifying variables which may predict the quality of a specific exercise.  The data is from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways, represented by the classe variable.

# Analysis

First, the appropriate R libraries and data need to be loaded.

```{r}
library(caret)
library(rpart)
trainingRaw <- read.csv("./pml-training.csv")
testing <- read.csv("./pml-testing.csv")
```

## Cleaning the Data

The training set contains many variables.

```{r}
dim(trainingRaw)
```

Getting rid of the variables which don't account for a significant level of the variance observed in the data will produce a more meaningful data set.

```{r}
nzv <- nearZeroVar(trainingRaw)
training01 <- trainingRaw[, -nzv]
```

A review of the new data set shows that the first five variables should have no predictive value. 

```{r}
head(trainingRaw[,1:5])
```

There are also some variables that show a large proportion of NA values.  The first five columns of the data will be dropped along with any variables that show a majority of NA values.

```{r}
training02 <- training01[, -(1:5)]
dimensions <- dim(training02)
rows <- dimensions[1]
training03 <- training02[ , colSums(is.na(training02)) < rows/2]
dim(training03)
```

A closer review of the data shows that this conveniently gets rid of all NA values.

```{r}
dim(training03[ , colSums(is.na(training03)) == 0])
```

## Building a Model

The data will now be partitioned to allow for cross-validation.  A seed value will also be set to allow for reproducibility.

```{r}
set.seed(12345)
inTrain <- createDataPartition(y=training03$classe, p=0.75, list=FALSE)
training <- training03[inTrain, ]
validation <- training03[-inTrain, ]
```

The data is now ready for modelling.  A decision tree may provide additional insight into the variables that have the most predictive value with respect to classe.  The following model uses recursive partitioning.

```{r}
modFit <- rpart(classe ~ ., method = "class", data = training)
plot(modFit, main = "Class Tree for Classe")
text(modFit)
```

## Prediction on Validation Set

The model may now be used on the validation set to identify the accuracy rate.

```{r}
predictions <- predict(modFit, validation, type = "class")
confusionMatrix(predictions, validation$classe)
```

The confusion matrix identifies an accuracy rate of 79%.

